Collective behavior of Janus Particles Suspended in a Viscous Fluid

### 2022-07-26 doc/PRF2022/PRF_main.tex
updated table/tidied section 2

\usepackage{todonotes}
\usepackage{graphicx}
%\usepackage{epstopdf,epsfig}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{commath}

\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newcommand{\RomanNumeralCaps}[1]
\linenumbers

@@

\subsection{Mobility problem}
%A suspension of $N_b$ Janus particles is immersed in a viscous solvent.
%RJR: In the above sentence, suspection and immersed are redundant 
The Janus particles are suspended in a viscous solvent.
The particles are disks of radius $c$, center $\aa_i$, and
orientation $\theta_i$ relative to the horizontal axis, where $i = 1, \dots, N_b$
and $N_b$ is the number of particles.
The domain $\Omega(t) \subset \mathbb{R}^2$ is the solvent phase and $t$
is time.  The boundary of
$\Omega$ is $\bd\Omega = \Gamma_1 \cup \cdots \cup \Gamma_{N_b}$, where
$\Gamma_i$ is the boundary of Janus particle $i$. Assuming inertial
terms are negligible, we have 
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity of the solvent, $p$ is the pressure,
$\uu_\infty$ is the background flow velocity, and $\mu$ is the constant
viscosity. The solvent velocity satisfies the no-slip boundary condition
for a rigid body motion
\begin{equation}
  \label{eq:rigid-vel}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, 
    \quad \xx \in \Gamma_i,
\end{equation}
where $\vv_i$ is the translational velocity, $\omega_i$ is the angular
velocity, and $\langle x, y \rangle^{\perp} = \langle -y, x\rangle$.
Without inertia, the system in~\eqref{eq:stokes}--\eqref{eq:rigid-vel}
also satisfies the force-free and torque-free conditions.
%This means
%that the total force $\FF_i,$ and torque $T_i$ from the hydrodynamic
%stress on each particle (with subscript `$i$') balance the force and
%torque from the hydrophobic potential~\cite{Fu2022_JFM}. 
The total force
and torque from the fluid flow are
\begin{align}
  \label{eq:force}
 \FF_i = \int_{\Gamma_i} \ssigma \cdot \nnu_i \, \dif s, \quad
 T_i = \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu_i) \, \dif s,
  \quad i=1,\ldots,N_b,
\end{align}
respectively, where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu +
\nabla \uu^T \right)$ is the hydrodynamic stress tensor (pressure
tensor), and $\nnu_i$ is the unit outward normal on the $i$th particle.

The total force $\FF_i,$ and torque $T_i$ from the hydrodynamic
stress on each particle (with subscript `$i$') balance the force and
torque from the hydrophobic potential. 
%$\FF_i$ and $T_i$ are to be in balance with the force and torque from the hydrophobic interactions between the
%particles and the viscous solvent. 
As in~\citet{Fu2022_JFM} such a hydrophobic attraction potential
$u(\xx,t)$ is assumed to the minimizer of the free energy
\begin{equation}
\label{eq:free_energy}
F = \gamma
\int_{\Omega} \rho |\nabla u|^2 + \rho^{-1} u^2
\,d\xx
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right), \quad
\end{equation}
where $u(\xx,t)$ is as an order parameter for the structure of water,
$\rho$ is a decay length, and $\gamma$ is an interfacial tension.  The
radius of the Janus particles $c$ also serves as an ``effective"
distance between particle centers, below which the steric repulsion
between a pair of JPs becomes important.

The governing equation for $u(\xx,t)$ is obtained by computing the
domain variation of the free energy $F$~\cite{Fu20}. We obtain a
screened Laplace equation boundary value problem as a model for the
hydrophobic interaction between water and JPs:
\begin{alignat}{3}
  \label{eq:SLpde}
  -\rho^2 \Delta u + u &= 0,\qquad &&\xx \in \Omega, \\
  \label{eq:SLbc}
  u &= g, &&\xx \in \bd\Omega, \\
  \label{eq:SLfarfield}
  u &\rightarrow 0, &&|\xx| \rightarrow \infty,
\end{alignat}
The boundary condition $g$ encodes hydrophobic properties of the
particle-solvent interface. The parameter $\rho_0$ is an ``effective"
distance between particle centers, below which the steric repulsion
between a pair of JPs becomes important.
\todo[inline]{We are calling both $c$ and $\rho_0$ as this ``effective"
distance (see above). Also, $\rho_0$ is not defined. Should it be
$\rho$?}

To calculate forces and torques from the free energy, we compute the
variation of the domain of the free energy~\eqref{eq:free_energy}
subject to~\eqref{eq:SLpde}--\eqref{eq:SLbc}, and obtain 
\begin{equation}
  \FF_i \equiv \int_{\Gamma_i} \mathbf{T}\nnu \, \dif s
  - \frac{M}{\rho_0}
  \sum_{j \neq i}
  \frac{\aa_i - \aa_j}{|\aa_i - \aa_j|}
P'\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right)
  ,\quad
T_i \equiv \int_{\Gamma_i} (\xx - \aa_i)^{\perp}\cdot (\mathbf{T}\nnu) \, \dif s, 
\end{equation}
for $i = 1,\dots,N_b$ with particle outward normal $\nu$. The term 
\begin{align}
\mathbf{T}
= \gamma \left[ \rho^{-1} u^2 \mathbf{I}
  + \rho \left(|\nabla u|^2 \mathbf{I} - 2\nabla u \nabla u^T\right)\right]
\end{align}
is the hydrophobic stress. Repulsion is rotationally symmetric and does
not contribute to the torque. The derivation details are found
in~\citet{Fu20}.

Since the force and torque from the hydrodynamic stress on each particle
are in balance with the total force and torque coming from the
hydrophobic potential and repulsion, we have
\begin{equation}
  \label{eq:force}
 \int_{\Gamma_i} \ssigma \cdot \nnu \, \dif s = \FF_i, \quad
 \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu) \, \dif s = T_i,
  \quad i=1,\ldots,N_b,
\end{equation}
where 
%$\FF_i \in \mathbb{R}^2$ and $T_i \in \mathbb{R}$
%are interparticle forces and torques. 
$\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor (pressure tensor).

The dimensionless repulsion profile $P$ takes the form $P(s) = 1 -
\sin(\pi s/2)$ for $0 \leq s < 1$ and $P(s) = 0$ for $1 \leq s <
\infty$.  As such, only those particles within a distance $\rho_0$ of
$\Gamma_i$ contribute to repulsion.  The parameter $M$ is the repulsion
modulus.

We use the parameters $\mu = 1$ cP, $\gamma = 4$ pN nm$^{-1}$,
$\rho = 5$ nm, $c = 1.25$ nm, $M = 2$ pN, and $\rho_0 = 0.5$ nm.

\subsection{Boundary integral represenation and time stepping}
We recast~\eqref{eq:stokes}--\eqref{eq:rigid-vel}
and~\eqref{eq:SLpde}--\eqref{eq:SLfarfield} as boundary integral
equations (BIEs) and discretize each BIE at $N$ points on each of the
$N_b$ particles with a collocation method. To express the solution
of~\eqref{eq:SLpde}--\eqref{eq:SLfarfield}, we adopt the double layer
potential
\begin{align}
\label{eq:SL_BIE}
  u({\xx}) = \DD[\sigma](\xx) = \int_\Gamma 
    \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, 
    \dif s_\yy, \quad \xx \in \Omega,
\end{align}
where $G(\xx) = \frac{1}{2\pi}K_0(|\xx|/\rho)$ is the fundamental
solution to the screened Laplace equation~\eqref{eq:SLpde}, $K_0$ is the
zeroth-order modified Bessel function of the first kind, and $\sigma$ is
a scalar-valued density function.  The subscript in $\dif s_\yy$ denotes
integration with respect to $\yy \in \Gamma$.  For the velocity, we use
the representation~\cite{pow-mir1987}
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega,
\end{align}
with the vector-valued density function $\eeta$, Stokeslet
$\SS(\xx,\aa_i) = \frac{1}{4\pi} \left(-\log |\rr|\mathbf{I} + \frac{\rr
\otimes \rr}{|\rr|^2}\right)$, and Rotlet $\RR(\xx,\aa_i) =
\frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2}$, $\rr = \xx - \aa_i$.

@@

In terms of fluid behavior,
the multilamellar (BC (ii)) structure behaves as shear thinning fluid
while the striated structure (BC (iii)) has a finite yield stress and
behaves as a Bingham plastic. 
Qualitatively, the strains for BC (ii) increases gradually 
with shear rate (Fig~\ref{fig:BC2_shear}(f)).
In contrast, the strains for BC (iii) are constant over a range of flow strengths
and then increase when the flow strengths are large enough
to deform the body (Fig~\ref{fig:BC3_shear}(f)).
To quantify this behavior, we form 
\[
\mu_{\text{sus}} = \frac{\dot \gamma}{\dot E} \mu
\]
as an effective dynamic viscosity of the suspension.
The numerator $\dot \gamma \mu$ gives the force per area provided by the solvent stresses. 
The strain rate $\dot E$ of the body is obtained by fitting the curves the strain curves e.g.,
Fig~\ref{fig:BC2_shear}(f).
Calculated suspension viscosities are shown
in Table \ref{tbl:bcii_visc}.
We observe that for the multilamellar structure, $\mu_{\text{sus}}$ is decreasing
with increasing shear rate.  The predicted suspension viscosities 
are on the order of hundreds of times that of water, consistent with
measurements of noninteracting particle suspensions\cite{KONIJN201461}.
In the case of striated structures, viscosity is effectively infinite
for the lower two of the shear rates, and then drops to a finite value $\mu_{\text{sus}} = 60$ mPa s
when the shear rate reaches $0.15$ ns$^{-1}$.
\begin{table}
  \caption{Viscosity $\mu_{\text{sus}}$ (mPa s) for the Multilamellar and Striated Suspensions    
\label{tbl:bcii_visc}
}
\centering
\begin{tabular}{c|c|c|c}
\multicolumn{4}{c}{multilamellar BC (ii)} \\
\hline
  $\dot \gamma$ & 0.05 & 0.10 & 0.15\\
  \hline
  $\phi$ & 0.50 & 0.47 & 0.41\\
  \hline
  $\mu_{\text{sus}} $ & 401 & 66 & 42\\
\hline
\end{tabular}
\quad \quad \quad
\begin{tabular}{c|c|c|c}
\multicolumn{4}{c}{striated BC (iii)} \\
\hline
  $\dot \gamma$ & 0.100 & 0.125 & 0.15\\
  \hline
  $\phi$ & 0.47 & 0.46 & 0.45\\
  \hline
  $\mu_{\text{sus}} $ & 2.5e3 & 1.4e3 & 60\\
\hline
\end{tabular}
\end{table}

>>>>

\usepackage{todonotes}
\usepackage{graphicx}
%\usepackage{epstopdf,epsfig}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{commath}
\usepackage{tabularx}
\usepackage{tabularray}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newcommand{\RomanNumeralCaps}[1]
\linenumbers

@@

\subsection{Mobility problem}
%A suspension of $N_b$ Janus particles is immersed in a viscous solvent.
%RJR: In the above sentence, suspection and immersed are redundant 
The Janus particles are suspended in a viscous solvent.
The particles are disks of radius $c$, center $\aa_i$, and
orientation $\theta_i$ relative to the horizontal axis, where $i = 1, \dots, N_b$
and $N_b$ is the number of particles.
The domain $\Omega(t) \subset \mathbb{R}^2$ is the solvent phase and $t$
is time.  The boundary of
$\Omega$ is $\bd\Omega = \Gamma_1 \cup \cdots \cup \Gamma_{N_b}$, where
$\Gamma_i$ is the boundary of Janus particle $i$. Assuming inertial
terms are negligible, we have 
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity of the solvent, $p$ is the pressure,
$\uu_\infty$ is the background flow velocity, and $\mu$ is the constant
viscosity. The solvent velocity satisfies the no-slip boundary condition
for a rigid body motion
\begin{equation}
  \label{eq:rigid-vel}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, 
    \quad \xx \in \Gamma_i,
\end{equation}
where $\vv_i$ is the translational velocity, $\omega_i$ is the angular
velocity, and $\langle x, y \rangle^{\perp} = \langle -y, x\rangle$.

Without inertia, $(\uu, p)$ also satisfy the force-free and torque-free conditions.
This means that the force and torque from the hydrodynamic stress on each particle 
are in balance with the total
force and torque coming from the hydrophobic potential and repulsion \cite{Fu2022_JFM}.
To compute this balance, we define
the free energy\cite{Fu2022_JFM}
\begin{equation}
\label{eq:free_energy}
F = \gamma
\int_{\Omega} \rho |\nabla u|^2 + \rho^{-1} u^2
\,d\xx
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right), \quad
\end{equation}
where $u(\xx,t)$ is as an order parameter for the structure of water,
$\rho$ is a decay length, and $\gamma$ is an interfacial tension.

The order parameter $u(\xx,t)$ is assumed to the minimizer of the free energy $F$ \cite{Fu20}.
Its governing equations $u(\xx,t)$ are a screened Laplace equation boundary value problem
\begin{gather}
  \label{eq:SL}
  -\rho^2 \Delta u + u = 0,\quad \xx \in \Omega, \\
  \label{eq:SLbc}
  u = g, \quad \xx \in \bd\Omega, \quad 
  u \rightarrow 0, \quad |\xx| \rightarrow \infty.
\end{gather}
The boundary condition $g$ encodes hydrophobic properties of the
particle-solvent interface.

The dimensionless repulsion profile $P$ takes the form $P(s) = 1 -
\sin(\pi s/2)$ for $0 \leq s < 1$ and $P(s) = 0$ for $1 \leq s <
\infty$. The parameter $\rho_0$ is an ``effective" distance between particle surfaces,
below which the steric repulsion between a pair of JPs becomes important.
The parameter $M$ is the repulsion modulus.

To get the force $\FF_i$ and torque $T_i$ on $\Gamma_i$ from the free energy,
we compute the variation of the domain of the free energy 
\eqref{eq:free_energy} subject to \eqref{eq:SL}-\eqref{eq:SLbc},
and obtain 
\begin{equation}
  \FF_i = \int_{\Gamma_i} \mathbf{T}\nnu \, \dif s
  - \frac{M}{\rho_0}
  \sum_{j \neq i}
  \frac{\aa_i - \aa_j}{|\aa_i - \aa_j|}
P'\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right)
  ,\quad
T_i = \int_{\Gamma_i} (\xx - \aa_i)^{\perp}\cdot (\mathbf{T}\nnu) \, \dif s, 
\end{equation}
for $i = 1,\dots,N_b$ with particle outward normal $\nu$. The term 
$
\mathbf{T}
= \gamma \left[ \rho^{-1} u^2 \mathbf{I}
  + \rho \left(|\nabla u|^2 \mathbf{I} - 2\nabla u \nabla u^T\right)\right]
$
is the hydrophobic stress.
Repulsion is rotationally symmetric and does not contribute to the torque.
The derivation details are found in \citet{Fu20}.

Since the force and torque from the hydrodynamic stress on each particle
are in balance with the total force and torque coming from the
hydrophobic potential and repulsion, we have
\begin{equation}
  \label{eq:force}
 \int_{\Gamma_i} \ssigma \cdot \nnu \, \dif s = \FF_i, \quad
 \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu) \, \dif s = T_i,
  \quad i=1,\ldots,N_b,
\end{equation}
where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor.

Once \eqref{eq:stokes}-\eqref{eq:force} have been solved,
the velocities $(\vv_i, \omega_i)$ from \eqref{eq:rigid-vel}
update the particle positions and orientations.
We use the parameters $\mu = 1$ cP, $\gamma = 4$ pN nm$^{-1}$,
$\rho = 5$ nm, $c = 1.25$ nm, $M = 2$ pN, and $\rho_0 = 0.5$ nm.
These values give a characteristic time $1$ ns and characteristic length $1$ nm.

\subsection{Boundary integral represenation and time stepping}
We recast \eqref{eq:stokes}--\eqref{eq:rigid-vel} and \eqref{eq:SL}
as boundary integral equations (BIEs) and discretize each BIE at
$N$ points on each of the $N_b$ particles with a collocation method.
To express the solution of \eqref{eq:SL},
we adopt the double layer potential
\begin{align}
\label{eq:SL_BIE}
u({\bf x}) = \DD[\sigma](\xx) = \int_\Gamma \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, \dif s_\yy
, \quad \xx \in \Omega,
\end{align}
where $G(\xx) = \frac{1}{2\pi}K_0(|\xx|/\rho)$ is the fundamental solution to the screened Laplace equation \eqref{eq:SL},
$K_0$ is the zeroth-order modified Bessel function of the first kind,
and $\sigma$ is a scalar-valued density function.
The subscript in $\dif s_\yy$ denotes integration with respect to $\yy \in \Gamma$. 
For the velocity, we use the representation [REF]
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega,
\end{align}
with the vector-valued density function $\eeta$,
Stokeslet 
$\SS(\xx,\aa_i) = \frac{1}{4\pi} \left(-\log |\rr|\mathbf{I} + \frac{\rr \otimes \rr}{|\rr|^2}\right)$,
and Rotlet
$\RR(\xx,\aa_i) = \frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2}$,
$\rr = \xx - \aa_i$ [REF].

@@

In terms of fluid behavior,
the multilamellar (BC (ii)) structure behaves as shear thinning fluid
while the striated structure (BC (iii)) has a finite yield stress and
behaves as a Bingham plastic. 
Qualitatively, the strains for BC (ii) increases gradually 
with shear rate (Fig~\ref{fig:BC2_shear}(f)).
In contrast, the strains for BC (iii) are constant over a range of flow strengths
and then increase when the flow strengths are large enough
to deform the body (Fig~\ref{fig:BC3_shear}(f)).
To quantify this behavior, we form 
\[
\mu_{\text{sus}} = \frac{\dot \gamma}{\dot E} \mu
\]
as an effective dynamic viscosity of the suspension.
The numerator $\dot \gamma \mu$ gives the force per area provided by the solvent stresses. 
The strain rate $\dot E$ of the body is the slope of a linear fit to the strain curves e.g.,
in Fig~\ref{fig:BC2_shear}(f).

Calculated suspension viscosities are shown
in Table \ref{tbl:bcii_visc}.
The middle row gives the solid area fraction $\phi$. 
We observe that for the multilamellar structure,
$\mu_{\text{sus}}$ is decreasing
with increasing shear rate.  The predicted suspension viscosities 
are on the order of hundreds of times that of water, consistent with
measurements of noninteracting particle suspensions\cite{KONIJN201461}.
In the case of striated structures, viscosity is effectively infinite
for the lower two of the shear rates, and then drops to a finite value $\mu_{\text{sus}} = 60$ mPa s
when the shear rate reaches $0.15$ ns$^{-1}$.
\begin{table}
  \caption{Suspension viscosity $\mu_{\text{sus}}$ (mPa s) 
\label{tbl:bcii_visc}
}
\centering
\begin{tabularx}{0.7\textwidth}{c|X|X|X||X|X|X}
&\multicolumn{3}{c||}{BC (ii)} & \multicolumn{3}{c}{BC (iii)}\\
\hline
  $\dot \gamma$ & 0.05 & 0.10 \quad & 0.15 & 0.100 & 0.125 & 0.15\\
  \hline
  $\phi$ & 0.50 & 0.47 & 0.41 & 0.47 & 0.46 & 0.45 \\
  \hline
  $\mu_{\text{sus}} $ & 401 & 66 & 42 & 2.5e3 & 1.4e3 & 60\\
\hline
\end{tabularx}
\end{table}
